SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python

### 2019-09-08 scipy-1.0/paper.tex
Update scipy-1.0/paper.tex
Co-Authored-By: Matt Haberland <mdhaber@mit.edu>

Python is an interpreted, high-level, general-purpose computer programming
language, designed by Guido van Rossum in the late 1980s,
with a dynamic type system and an emphasis on readability and rapid prototyping\cite{cpython-refman,cpython-source}.
% https://legacy.python.org/search/hypermail/python-1992/0000.html
%
As a general purpose programming language, it had no special support for
scientific data structures or algorithms, unlike many of the other established
computation platforms of the time. Yet, scientists soon discovered the
language's virtues, such as its ability to wrap C and Fortran
libraries and to then drive those libraries interactively.  Scientists
could thereby gain access to a wide variety of existing computational
libraries without concerning themselves with low-level programming
concepts such as memory management.

@@

    Special Functions (Bessel, Hankel, Airy, etc.) % hanker -> Hankel
    Signal/Image Processing
    2D Plotting capabilities
    Integration
    ODE solvers
    Optimization (simplex, BFGS, Newton-CG, etc.) % Netwon -> Newton
    Genetic Algorithms
    Numeric -> C++ expression compiler
    Parallel programming tools
    Splines and Interpolation
    And other stuff.
\end{verbatim}
\caption{Excerpt from the SciPy 0.1 release announcement (typos corrected) posted August 20, 2001, on the ``Python-list'' mailing list\cite{first-scipy-rel}.}\label{fig:announce-0.1}
\end{figure}

@@

By the middle to late 2000s, SciPy was starting to mature after a long phase of significant
growth and adoption. The scope of the SciPy library
narrowed, while the breadth of the ecosystem grew
through a new type of auxiliary package: the
scikit\cite{scikits-general}, a complementary library
developed outside of SciPy, allowing for more rapid exploration of
experimental ideas while maintaining familiar style and development methodology.
In SciPy itself, tooling, development, documentation, and release processes became more professional.
The library was expanded carefully, with the patience affordable in open source
projects and via best practices common in industry \cite{millman2014developing}.

@@

\subsection*{SciPy Today}
At the time of writing, the SciPy library consists of nearly
\num{600000}~lines of code organized in 16~subpackages.
Over \num{110000}~GitHub repositories and \num{6500}~packages depend
on SciPy\cite{dependents}. Some of the major
feature highlights from the three years preceding
SciPy 1.0 are discussed in the Key Technical Improvements section below,
and milestones in its history are highlighted in Figure~\ref{fig:timeline}.
% At the time of writing, SciPy produces a new release
% approximately every six months, and some of the major
% feature highlights from the three years preceding
% SciPy 1.0 are discussed in a technical improvements
% section below. Some longer-term milestone events are
% highlighted in Figure~\ref{fig:timeline}.

@@

Fortran, despite its age, is still a high-performance scientific programming language with
continued contemporary usage\cite{Koelbel:1993:HPF:562354}. Thus, we wrap the following excellent, field-tested Fortran
libraries in order to provide Python convenience while benefiting from their performance:
QUADPACK\cite{piessens1983quadpack} and ODEPACK\cite{hindmarsh1983odepack} for numerical integration
and solution of initial value problems;
FITPACK\cite{Dierckx:1993:CSF:151103}, ODRPACK\cite{ODRPACK_Boggs}, and MINPACK\cite{osti_6997568} for
curve-fitting and least-squares minimization;
FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251} for performing Fourier transforms;
ARPACK\cite{leh:sor:yan96} for solving eigenvalue problems;
ALGORITHM 644\cite{Amos:1986:APP:7921.214331} for computing Bessel Functions; and
CDFLIB\cite{CDFLIB_site} for evaluating cumulative density functions.

@@

Test-driven development has been described as a way to manage fear and
uncertainity when making code changes\cite{TDD-Beck}. For each component of SciPy, we write multiple small executable
tests that verify its intended behavior. The collection of these, known as a \emph{test suite}, increases confidence in
the correctness and accuracy of the library, and lets us make code modifications known not to alter desired behavior. Monitoring the number of lines
of code in SciPy covered by unit tests is one way we maintain
some certainty that changes and new features are correctly implemented.
The SciPy test suite is orchestrated by a continuous integration matrix that
includes POSIX and Windows (32/64-bit) platforms managed by Travis CI and
AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There
are more than \num{13000}~unit tests in the test suite, which is written for usage
with the \texttt{pytest}\cite{pytest} framework.
Test coverage at the SciPy~1.0 release point (Figure~\ref{fig:coverage}) was at
87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov} and
45\% for compiled (C, C++, and Fortran) code according to \texttt{gcov}\cite{gcov}.
Documentation for the code is automatically built and published by
the CircleCI service to facilitate evaluation of documentation changes /
integrity.  Our full test suite also passes with PyPy3\cite{Bolz:2009:TMP:1565824.1565827}, a just-in-time compiled
version of the Python language.

>>>>

Python is an interpreted, high-level, general-purpose computer programming
language, designed by Guido van Rossum in the late 1980s,
with a dynamic type system and an emphasis on readability and rapid prototyping.
The reference and most popular implementation of Python is
CPython\cite{cpython-refman,cpython-source}, which is written
in the C and Python languages and assumed throughout this paper.
% https://legacy.python.org/search/hypermail/python-1992/0000.html
%
As a general purpose programming language, it had no special support for
scientific data structures or algorithms, unlike many of the other established
computation platforms of the time. Yet, scientists soon discovered the
language's virtues, such as its ability to wrap C and Fortran
libraries and to then drive those libraries interactively.  Scientists
could thereby gain access to a wide variety of existing computational
libraries without concerning themselves with low-level programming
concepts such as memory management.

@@

    Special Functions (Bessel, Hankel, Airy, etc.) % hanker -> Hankel
    Signal/Image Processing
    2D Plotting capabilities
    Integration
    ODE solvers
    Optimization (simplex, BFGS, Newton-CG, etc.) % Netwon -> Newton
    Genetic Algorithms
    Numeric -> C++ expression compiler
    Parallel programming tools
    Splines and Interpolation
    And other stuff.
\end{verbatim}
\caption{Excerpt from SciPy 0.1 release announcement (typos corrected).}\label{fig:announce-0.1}
\end{figure}

@@

By the middle to late 2000s, SciPy was starting to mature after a long phase of significant
growth and adoption. The scope of the SciPy library
narrowed, while the breadth of the ecosystem grew
through a new type of auxiliary package: the scikit\cite{scikits-general}.
The tooling, development, documentation, and release processes became more professional.
SciPy was expanded carefully, with the patience affordable in open source
projects and via best practices common in industry \cite{millman2014developing}.

@@

\subsection*{SciPy Today}
At the time of writing, the SciPy library consists of nearly
\num{600000}~lines of code organized in 16~subpackages. The development
team and community currently interact and operate primarily
on GitHub, an online version control and task management platform.
Over \num{110000}~GitHub repositories and \num{6500}~packages depend
on SciPy\cite{dependents}. Some of the major
feature highlights from the three years preceding
SciPy 1.0 are discussed in the Key Technical Improvements section below,
and milestones in its history are highlighted in Figure~\ref{fig:timeline}.
% At the time of writing, SciPy produces a new release
% approximately every six months, and some of the major
% feature highlights from the three years preceding
% SciPy 1.0 are discussed in a technical improvements
% section below. Some longer-term milestone events are
% highlighted in Figure~\ref{fig:timeline}.

@@

Fortran, despite its age, is still a high-performance scientific programming language with
continued contemporary usage\cite{Koelbel:1993:HPF:562354}. Thus, we wrap the following excellent, field-tested Fortran
libraries in order to provide Python convenience while benefiting from their performance:
FFTPACK\cite{SWARZTRAUBER198445, SWARZTRAUBER198251}, % for performing Fourier transforms,
ODEPACK\cite{hindmarsh1983odepack}, % for solving ordinary differential equation initial value problems,
QUADPACK\cite{piessens1983quadpack}, % for numerical integration of one-dimensional functions,
FITPACK\cite{Dierckx:1993:CSF:151103}, % for curve-fitting and interpolation,
ODRPACK\cite{ODRPACK_Boggs}, % for orthogonal distance regression,
MINPACK\cite{osti_6997568}, % for minimization of linear and nonlinear equations,
ARPACK\cite{leh:sor:yan96}, % for solving large scale eigenvalue problems,
ALGORITHM 644\cite{Amos:1986:APP:7921.214331}, and % for Bessel Functions, and
CDFLIB\cite{CDFLIB_site}. % for evaluation of cumulative density functions.

@@

The SciPy test suite is orchestrated by a continuous integration matrix that
includes POSIX and Windows (32/64-bit) platforms managed by Travis CI and
AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There
are more than \num{13000}~unit tests in the test suite, which is written for usage
with the \texttt{pytest}\cite{pytest} framework.
Test coverage at the SciPy~1.0 release point (Figure~\ref{fig:coverage}) was at
87\% for Python code according to \texttt{pytest-cov}\cite{pytest-cov} and
45\% for compiled (C, C++, and Fortran) code according to \texttt{gcov}\cite{gcov}.
Documentation for the code is automatically built and published by
the CircleCI service to facilitate evaluation of documentation changes /
integrity.  Our full test suite also passes with PyPy3\cite{Bolz:2009:TMP:1565824.1565827}, a just-in-time compiled
version of the Python language.
